--- 
site: bookdown::bookdown_site

author: "peacecop kalmer:"
colorlinks: TRUE
date: "`r Sys.Date()`"
description: "Tallinna Ülikooli akustika muusikas kursuse peacecop kalmeri: raport"
lang: et
include-before:
    - \input{title-page.tex}

documentclass: book
papersize: a4
geometry: top=2cm, right=2cm, bottom=3cm, left=3cm
classoption: onecolumn, twoside
linestretch: 1.5

fontsize: 12pt
mainfont: DejaVu Serif # for displaying both Chinese symbols and emojis
CJKmainfont: Droid Sans Fallback

bibliography: [references.bib]
csl: apa-numeric-superscript-brackets.csl
link-citations: yes

lof: yes
lot: yes

---

```{r setup, echo=FALSE, message=FALSE}
# example chunk options set globally
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE
)

local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- strrep(' ', max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})

options(OutDec = ",")

if (!require("librarian")) {
  install.packages("librarian", dependencies = TRUE)
}

library(librarian)

#' This function displays external graphics by making the appropriate conversion aforehead.
#' 
#' @param path the path to the image file
include_external_graphics = function(path) {

  if (knitr::is_latex_output() && ("pdf" != xfun::file_ext(path))) {
    output = xfun::with_ext(path, 'pdf')
    
    if ("svg" == xfun::file_ext(path)) {
      system2('rsvg-convert', c('-f', 'pdf', '-a', '-o', shQuote(c(output, path))))
    } else {
      system(paste('gm convert ', path, ' ', output), intern = FALSE, ignore.stdout = FALSE, ignore.stderr = FALSE, wait = TRUE)
    }
    
  } else {
    output = path
  }
  
  knitr::include_graphics(output)
}

#' This function creates and displays a table.
#' 
#' @param table the data frame or matrix
#' @param caption the caption
#' @param digits the number of significant digits
#' @param additional_header the additional header
#' @param footnotes the array of footnotes
#' @param do_i_transpose Do i transpose?
#' 
#' @returns the table
print_table <- function(table, caption, digits = 2, additional_header = NULL, footnotes = NULL, do_i_transpose = FALSE) {

  if (!is.null(knitr::opts_knit$get('rmarkdown.pandoc.to')) && (knitr::opts_knit$get('rmarkdown.pandoc.to') == "docx")) {
    librarian::shelf("flextable")
    table <- (flextable(table))
  } else {
    librarian::shelf("kableExtra")

    if (do_i_transpose) {
      transposed_table <- t(table)
      rownames(transposed_table) <- colnames(table)
      colnames(transposed_table) <- rownames(table)
      table <- transposed_table
    }
    
    table <- kable(table, booktabs = TRUE, escape = FALSE, caption = caption, digits = digits, longtable = TRUE)

    if (!is.null(additional_header)) {
      table <- add_header_above(kable_input = table, header = additional_header, escape = FALSE)
    }
    
    table <- (kable_styling(
      table,
      bootstrap_options = c("striped", "bordered", "condensed", "responsive", "hover"),
      latex_options = c("striped", "repeat_header"),
      # repeat_header_text = paste(caption, "$\\textit{(Jätkub...)}$"),
      repeat_header_method = "replace",
      repeat_header_continued = TRUE # "Tabel järgneb järgmisel leheküljel..."
    ))
    
    if (!is.null(footnotes)) {
      table <- footnote(kable_input = table, number = footnotes, escape = FALSE, threeparttable = TRUE)
    }
    
  } 

  return(table)
}

#' This function prints the non-math LaTeX text no matter the output mode is.
#'
#' @param text the LaTeX text where every command must have doublebackslash in front
#' 
#' @example print_latex("\\LaTeX")
#' 
#' @returns the formatted string that can be used for any output
render_nonmath_LaTeX_string_universal <- function(text) {
  
  if (knitr::is_html_output()) {
    formatted_string <- paste("$\\mathrm{", text, "}$")
  } else {
    formatted_string <- text
  }

  return(formatted_string)
}

#' This function renders the given text bold no matter what the output mode is. The function is needed if the main font has no ability to create bold symbols.
#' 
#' @param text the text to be rendered bold
#' @param to_italic Do we render into italic? Or rather bold?
#' 
#' @returns the text in italic or bold
render_font <- function(text, to_italic = TRUE) {
  
  if (knitr::is_html_output()) {
    
    if (to_italic) {
      formatted_string <- paste("*", text, "*", sep = "")
    } else {
      formatted_string <- paste("**", text, "**", sep = "")
    }
    
  } else {
    
    if (to_italic) {
      formatted_string <- paste("\\fontspec{DejaVu Sans}{\\textit{", text, "}}", sep = "")
    } else {
      formatted_string <- paste("\\fontspec{DejaVu Sans}{\\textbf{", text, "}}", sep = "")
    }
    
  }

  return(formatted_string)
}

#' This function renders the given text having emojis correctly because it uses a font that supports them. That font can't display anything in bold nor italic.
#' 
#' @param text the given text that contains emojis
#' 
#' @returns the input text with the emoji font
render_with_emojis <- function(text) {
  return(paste("\\fontspec{Symbola}", text, "\\fontspec{DejaVu Serif}{}", sep = ""))
}

#' This function renders a reference to a page as it cannot be used in the web output.
#' 
#' @param reference the reference identifier
#' @param escape Do we need to double-escape because we call the function from a table generation function?
#' 
#' @returns the rendered page reference
render_pageref <- function(reference, escape = FALSE) {
  
  if (knitr::is_html_output()) {
    rendered_pageref <- ""
  } else {
    
    if (escape) {
      rendered_pageref <- paste(" $\\\\pageref{", reference, "}$", sep = "")
    } else {
      rendered_pageref <- paste(" $\\pageref{", reference, "}$", sep = "")
    }
  }

  return(rendered_pageref)
}

#' This function colors text.
#' 
#' @param color the color code in hexadecimals
#' @param text the text to be colored
#' 
#' @returns the colored text
render_color <- function(color, text) {
  
  if (knitr::is_html_output()) {
    rendered_color <- paste("$\\color{#", color, "}{", text, "}$", sep = "")
  } else {
    rendered_color <- paste("$\\textcolor[HTML]{", color, "}{", text, "}$", sep = "")
  }

  return(rendered_color)
}

```

# Reeglistik

```{r}
I_0 <- 1e-12

```

Inimese kuulmislävel on võimsustihedus $\qty{`r as.character(I_0)`}{\W\per\m\squared}$ [@haynes_2014_crc(lk 14-52)].

Helitaseme arvutamiseks võimsustiheduste järgi saab kasutada valemit \@ref(eq:sound-level) leheküljel `r render_pageref(reference = "eq:sound-level")`. Selle valemi dimensionaalanalüüs on esitatud avaldises \@ref(eq:dimensional-analysis-of-sound-level) leheküljel `r render_pageref(reference = "eq:dimensional-analysis-of-sound-level")`, millest johtuvalt on ühik 1 või spetsiifiliselt detsibell.

```{=tex}
\begin{align}
\beta = 10 \times \mathrm{log}(\frac{I}{\qty{`r as.character(I_0)`}{\W\per\m\squared}})
(\#eq:sound-level)
\end{align}

```

[@giancoli_2014_physics(lk 331)].

```{=tex}
\begin{align}
1 \times \mathrm{log}(\frac{\mathrm{\frac{M \times L \times L}{T^3 \times L^2}}}{\mathrm{\frac{M \times L \times L}{T^3 \times L^2}}}).
(\#eq:dimensional-analysis-of-sound-level)
\end{align}

```

```{r label = "calculate-sound-level"}
#' This function calculates the sound level.
#' 
#' @param I the power intensity
#' 
#' @returns the sound level
calculate_sound_level <- function(I) {
  return(10 * log10(I / I_0))
}

```

Avaldan helitaseme valemist \@ref(eq:sound-level) leheküljel `r render_pageref(reference = "eq:sound-level")` võimsusintensiivsuse avaldises \@ref(eq:power-intensity) leheküljel `r render_pageref(reference = "eq:power-intensity")`

```{=tex}
\begin{align}
\frac{\beta}{10} &= \mathrm{log}(\frac{I}{\qty{`r as.character(I_0)`}{\W\per\m\squared}}) \\
10^{\frac{\beta}{10}} &= \frac{I}{\qty{`r as.character(I_0)`}{\W\per\m\squared}} \\
I &= \qty{`r as.character(I_0)`}{\W\per\m\squared} \times 10^{\frac{\beta}{10}}.
(\#eq:power-intensity)
\end{align}

```

```{r label = "calculate-power-intensity"}
#' This function calculates the power intensity.
#' 
#' @param beta the sound level
#' 
#' @returns the power intensity
calculate_power_intensity <- function(beta) {
  return(I_0 * 10^(beta / 10))
}

```

Võimsusintensiivsuse sõltuvus kaugusest on esitatud avaldises \@ref(eq:dependence-of-power-intensity-on-distance) leheküljel `r render_pageref(reference = "eq:dependence-of-power-intensity-on-distance")`.

```{=tex}
\begin{align}
I = \frac{P}{4 \times \pi \times r^2},
(\#eq:dependence-of-power-intensity-on-distance)
\end{align}

```

milles $P$ on võimsus ja $r$ kaugus heliallikast [@giancoli_2014_physics(lk 310)].
